### 2021-10-28 text/outline/outline.tex
add assay overview text

The closest parallels to our work in the statistical methodology literature are Gr\"{u}n \& Leisch \cite{Grun2008} and Ibrahim \cite{Ibrahim1990}. Gr\"{u}n \& Leisch derived a method for estimation and inference in a $k$-component mixture of GLMs. While we prefer to view GLM-EIV as a generalized errors-in-variables method,  the GLM-EIV model is equivalent to a two-component mixture of \textit{products} of GLM densities. Ibrahim proposed a procedure for fitting GLMs in the presence of missing-at-random covariates. Our method, by contrast, involves fitting two conditionally independent GLMs in the presence of a totally latent covariate. Thus, while Ibrahim and Gr\"{u}n \& Leisch are helpful references, our estimation and inference tasks are more complex.

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a separate work we proposed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and providing robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}, which, like SCEPTRE, focus on hypothesis testing, but unlike SCEPTRE, ignore the countedness of the data and do not adjust for confounders. Our focus in this work are on challenges complimentary to those that SCEPTRE addresses; most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression, moving beyond hypothesis testing.

\subsection{Assay overview}

\subsection{Analysis challenges}

Our focus in this work is on high multiplicity-of-infection (MOI), enhancer-targeting, single-cell CRISPR screens. In this section we cover relevant biological background and motivation.
\\ \\
\textbf{P1}: The human genome consists of genes (segments of DNA that code for proteins), enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic regions. Genome-wide association studies have revealed that the majority ($>95\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers. These noncoding variants contribute to disease by modulating the expression one or more genes, which in turn encode proteins that affect the phenotype. A central open challenge in genetics, therefore, is to link enhancers that harbor disease-associated variants to the genes that they target at genome-wide scale.
\\ \\
\textbf{P2}: High MOI single-cell CRISPR screens are the most promising biotechnology for solving this problem. \blue{Describe the experimental protocol here. Explain that we use the terms ``barcodes'' and ``gRNAs'' interchangeably, as polyadenylated gRNAs serve as barcodes in CROP-seq. Link to Figure \ref{analysis_challenges}.}
\\ \\
\textbf{P3}: Single-cell CRISPR screens pose several core analysis challenges. \blue{Describe the analysis challenges here: (i) unobserved perturbation; (ii) existence of background reads; (iii) highly discrete count data; (iv) nuisance variables}.

>>>>

The closest parallels to our work in the statistical methodology literature are Gr\"{u}n \& Leisch \cite{Grun2008} and Ibrahim \cite{Ibrahim1990}. Gr\"{u}n \& Leisch derived a method for estimation and inference in a $k$-component mixture of GLMs. While we prefer to view GLM-EIV as a generalized errors-in-variables method,  the GLM-EIV model is equivalent to a two-component mixture of \textit{products} of GLM densities. Ibrahim proposed a procedure for fitting GLMs in the presence of missing-at-random covariates. Our method, by contrast, involves fitting two conditionally independent GLMs in the presence of a totally latent covariate. Thus, while Ibrahim and Gr\"{u}n \& Leisch are helpful references, our estimation and inference tasks are more complex than theirs.

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. In a prior work we developed SCEPTRE \cite{Barry2020}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. Other applied methods for single-cell CRISPR screen analysis include MIMOSCA \cite{Dixit2016} and scMAGeCK \cite{Yang2019}. These methods, like SCEPTRE, focus on hypothesis testing (rather than estimation), but unlike SCEPTRE, they ignore the countedness of the data and are unable to handle confounders. In this work we tackle a set of analysis challenges that are complimentary to the challenges addressed by SCEPTRE; most importantly, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, moving beyond hypothesis testing.

\subsection{Assay overview and analysis challenges}

There are several broad classes of single-cell CRISPR screen assays, each suited to answer a different set of biological questions \cite{Gasperini2019,Datlinger2021,Mimitou2019}. In this work we focus on so-called high-multiplicity of infection (MOI) single-cell CRISPR screens. We expect the ideas that we develop for this assay to apply (with some effort) to other classes of single-cell CRISPR screens as well. This section motivates high MOI single-cell screens and provides some background on the experimental protocol and relevant analysis challenges.

The human genome consists of genes, enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic elements (that are not of importance to the current discussion). Genome-wide association studies (GWAS) have revealed that the majority ($>90\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers \cite{Gallagher2018}. These noncoding variants are thought to contribute to disease by modulating the expression one or more disease-relevant genes. Unfortunately, we do not know which gene (or genes) through which most noncoding variants exert their effect, limiting the interpretability of GWAS results. A central open challenge in genetics, therefore, is to link enhancers that harbor GWAS variants to the genes that they target at genome-wide scale \cite{Gasperini2020}.

The most promising biotechnology for solving this challenge are high MOI single-cell CRISPR screens. High MOI single-cell CRISPR screens combine CRISPR interference (CRISPRi) -- a variant of CRISPR that \textit{represses} a targeted region of the genome, leaving the primary DNA sequence intact  -- with single-cell sequencing. The experimental protocol, at a high level, is as follows. First, the scientist designs a library of several hundred to several thousand CRISPRi perturbations, each targeting a candidate enhancer for repression. Next, the scientist cultures several hundred thousand cells and delivers the CRISPRi perturbations to these cells. The perturbations assort into the cells randomly, with each cell receiving on average 10-40 distinct perturbations.

After waiting several days for CRISPRi to take effect, the scientist profiles each cell's transcriptome (i.e., its set of gene expressions) and perturbations. Finally, the scientist conducts perturbation-to-gene association analyses. Figure \ref{analysis_challenges}a depicts this process schematically, with colored bars (blue, red, and purple) representing distinct perturbations. For a given perturbation (e.g., the perturbation represented in blue), the scientist partitions the cells into two sets: those that received the perturbation (top) and those that did not (bottom). Next, for a given gene, the scientist runs a differential expression analysis across the two groups of cells, producing an estimate for the change in expression in response to the perturbation. If the estimated change in expression is large, the scientist can conclude that the enhancer targeted by the perturbation exerts a strong regulatory effect on the gene. This procedure is repeated for a large set of preselected perturbation-gene pairs.

% the scientist carries out the standard single-cell sequencing procedure to assay each cell's transcriptome -- -- and

% Discuss bardodes and gRNAs in the subsequent paragraph.
% \textbf{P3}: Single-cell CRISPR screens pose several core analysis challenges. \blue{Describe the analysis challenges here: (i) unobserved perturbation; (ii) existence of background reads; (iii) highly discrete count data; (iv) nuisance variables}.
